---
title: "Selling SPY Weeklys: 2014 - 2018"
subtitle: "A Simple Approach"
author: "Pritam Dalal"
date: "December 31, 2018"
output: 
  html_document:
    theme: united
    df_print: kable
editor_options: 
  chunk_output_type: console
---

<style>
    body .main-container {
        max-width: 700px;
    }
</style>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, load_packages, include = FALSE}
# loading packages
library(tidyverse)
library(tidyquant)
library(bizdays)
load_rmetrics_calendars(2000:2020)
bizdays.options$set(default.calendar="Rmetrics/NYSE")
```

```{r, read_data, include = FALSE}
###################
# reading in data #
###################

df_chain_desc <- 
    read_csv("/Users/Pritam/files/ods/plx_2018/data_output/spy_weekly_chain_desc_5yr.csv", col_types = cols())
df_spy_weekly_pnl_all <- 
    read_csv("/Users/Pritam/files/ods/plx_2018/data_output/spy_weekly_pnl_strangle_5yr.csv", col_type = cols())

df_spy_weekly_position_size_all <- 
    read_csv("/Users/Pritam/files/ods/plx_2018/data_output/spy_weekly_position_scaling_5yr.csv", col_type = cols())


```


```{r, include=FALSE}
dt_start <- min(df_chain_desc$execution)
dt_end <- max(df_chain_desc$last_trade_date)
int_trading_days <- bizdays(dt_start, dt_end)
dbl_aum <- 100000
dbl_base_size <- 75
dbl_aum_mult <- dbl_aum / dbl_base_size
```




```{r, combine_dataframes, include = FALSE}
####################################
## combining dataframes as needed ##
####################################
df_pnl_all <- df_spy_weekly_pnl_all

df_position_size_all <- df_spy_weekly_position_size_all

```




```{r, filtering_data, include = FALSE}
############################################
## filtering and adding columns as needed ##
############################################
# position size - filtering by underlying and variation
df_position_size <- 
    df_position_size_all %>% 
    filter(underlying == "SPY") %>%
    filter(variation == 0.1)

    
# pnls - scaling the pnl and rearranging columns
df_pnl <-         
    df_pnl_all %>% 
    filter(underlying_symbol == "SPY") %>%
    filter(variation == 0.1) %>% 
    mutate(
        naked_dly  = dly_opt_pnl * strangle_mult * dbl_aum_mult
        , dh_dly = dly_tot_pnl * strangle_mult * dbl_aum_mult
    ) %>% 
    select(underlying_symbol:strangle_mult, naked_dly:dh_dly) %>% 
    arrange(underlying_symbol, type, expiration, data_date)
```

```{r, spy_performance, include = FALSE}
# getting SPY data to throw in the comparison, using tidyquant package
df_spy_px <- 
    tq_get("SPY", get = "stock.prices", from = "2013-12-20", to = "2018-11-30")
# calculating log returns
df_spy_px$ret <- 
    log(df_spy_px$adjusted) - log(lag(df_spy_px$adjusted, 1)) 
# calculating PNL
df_spy_px <-
    df_spy_px %>% 
        mutate(
            pnl = dbl_aum * ret
        ) 
```

<!-- The purpose of this analysis is to demonstrate the capabilities of the Option Data Science (ODS) backtesting framework.  Towards this end, we will examine four different SPY option premium selling strategies.  Specifically, we consider the repeated selling of weekly and monthly options, both naked and delta-hedged.  The report ends with a proposal for a more involved backtest - the proposal includes deliverables, timeline, and pricing. -->

The purpose of this analysis is to backtest a simple strangle selling strategy on SPY weekly options over the last 5 years.  This simple trading strategy is a naive version of the one that ODS is analyzing for Parallax.

This report serves as something of a status update for the Parallax backtesting project.


#### Backtesting Parameters

**Dates:** 12/20/2013 - 11/30/2018 (`r int_trading_days` total trading days)

**Underlying:** SPY

**AUM:** $100,000

**Trades:** On each Friday, sell 10-delta strangles that expire the following Friday.


**Hedging Strategies:** Both naked and daily delta-hedged are considered.

<!-- **Monthly:** on each regular option expiration (3rd Friday of the month), sell 10-delta strangles that expire on the next regular expiration. -->


<!-- For both tenors, the first trades open on 12/16/2016 and the last trades close on 7/20/2018.  The monthly strategy is executed over 19 consecutive regular expirations, while the weekly strategy is executed over 83 consecutive weekly expirations.   -->


#### Premium 
Since we are selling options, let's consider the total the amount of premium sold and total number of options sold.  Here is a table of summary statistics: 
```{r, und_exp_prem_sold, echo = FALSE, , fig.cap="Premium Related Statistics"}
df_position_size %>% 
    #group_by(frequency) %>% 
    summarize(
        tot_prem = (sum(strangle_prem_sold) * dbl_aum_mult) %>% round(0)
        , avg_prem = (mean(strangle_prem_sold) * dbl_aum_mult) %>% round(0)
        , tot_opt = (sum(strangle_mult) * dbl_aum_mult / 100) %>% round(0)
        , avg_opt = (mean(strangle_mult) * dbl_aum_mult / 100) %>% round(0)
    )
    #.[c(2, 1), ]
```

Notice that we sell just under \$80K in premium over the backtest period: $310 per execution, and a total of 256 executions.  Each week we sell 10 strangles.


#### Performance
The table below displays performance of the two hedging strategies, as well as the same size investment in the S&P 500.
```{r, gathering_pnl, include = FALSE}
########################################################
## gathered the PNLs together and performing analysis ##
########################################################
# daily pnl for each strategy, all underlyings combined
df_dly_pnl <-
    df_pnl %>%
        group_by(data_date) %>%
        arrange(data_date) %>%
        summarize(
            naked_dly = sum(naked_dly)
            , dh_dly = sum(dh_dly)
        )

# gathering the dh_pnl column
df_pnl_gathered <-  
    df_dly_pnl %>% 
        rename(naked = naked_dly, delta_hedge = dh_dly) %>% 
        gather(hedge, daily_pnl, naked:delta_hedge) %>% 
        # mutate(strategy = paste0(frequency, "_", hedge)) %>% 
        # ungroup() %>% 
        select(strategy = hedge, data_date, daily_pnl) %>% 
        bind_rows(
            df_spy_px %>% 
                mutate(
                    strategy = "spy"
                    , daily_pnl = replace_na(pnl, 0)
                ) %>%
                select(strategy, data_date = date, daily_pnl)
        ) 


# calculating trade-to-date PNLs
df_pnl_gathered <- 
    df_pnl_gathered %>% 
        group_by(strategy) %>% 
        mutate(
            ttd_pnl = with_order(data_date, cumsum, daily_pnl)
        ) %>% 
        mutate(
            draw_down = cummax(ttd_pnl) - ttd_pnl
        )


# calculating various performance metrics
df_performance <- 
    df_pnl_gathered %>% 
        #filter(strategy != "spy") %>% 
        group_by(strategy) %>% 
        summarize(
            tot_pnl = sum(daily_pnl) %>% round(0)
            , sharpe_ratio = 
                ((mean(daily_pnl) / sd(daily_pnl)) * sqrt(252) )%>% round(2)
            , ret = 
                ((((sum(daily_pnl) + dbl_aum) / dbl_aum) ^ (252/int_trading_days)) - 1) %>% 
                round(3)
            , max_drawdown = (max(draw_down) / dbl_aum) %>% round(2)
        ) #%>% 
        #.[c(3, 4, 1, 2), ]

```


```{r, performance_table, echo = FALSE}
df_performance
```

Let's parse the performance of the delta-hedged strategy in greater detail.  Of the \$80K of premium sold, $22K is retained, which equates to an annualized return of 4.2%; the Sharpe-Ratio for this strategy was 1.31.  Compare this to investing in the S&P 500 which had returns of 8.6% and an SR of 0.80.  

Note: this PNL does not account for option commissions.


#### Visualizing Cumulative Performance

The the trade-to-date PNL graphs of these strategies are representative of the nature of naive option selling.  Everything is great...until it isn't.

```{r, ttd_pnl_graph, echo = FALSE}
# df_pnl %>% 
#     group_by(data_date) %>%
#     arrange(data_date) %>% 
#     summarize(
#         naked_dly = sum(naked_dly)
#         , dh_dly = sum(dh_dly) 
#     ) %>% 
#     mutate(
#         naked_ttd = cumsum(naked_dly)
#         , dh_ttd = cumsum(dh_dly)
#     ) %>% 
#     ggplot() +
#         geom_line(aes(x = data_date, y = naked_ttd), color = "green") +
#         geom_line(aes(x = data_date, y = dh_ttd), color = "blue")

df_pnl_gathered %>% 
    #filter(strategy != "spy") %>% 
    ggplot() +
    geom_line(aes(x = data_date, y = ttd_pnl, color = strategy)) +
    labs(
        title = "Selling SPY Strangles - Cumulative PNL"
        , subtitle = "$100K aum; 10-delta; Dec 20 2013 - Nov 30 2018"
        , x = "trade date"
        , y = "ttd pnl"
        #, caption =  "option prices from Delta-Neutral"

    )

```


#### PNL by Year

If we examine the PNL by year, we see that the worst year for option selling was 2018.  During the backtest period, the SP&500 significantly outperforms option selling in all years except for 2015.
```{r, pnl_by_year,echo = FALSE}

#df_pnl_gathered %>% filter(strategy == "spy") %>% View()

df_pnl_gathered %>%
    filter(data_date >= lubridate::ymd(20140101)) %>% 
    filter(!is.na(daily_pnl)) %>% 
    group_by(strategy, year = lubridate::year(data_date)) %>%
    summarize(
        pnl = sum(daily_pnl)
    ) %>% 
    ungroup() %>%
    ggplot(aes(factor(year), pnl, fill = strategy)) +
    geom_bar(stat = "identity", position = "dodge") +
    labs(
        title = "PNL by Year"
        , subtitle = "all strategies"
        , x = "trade date"
        , y = "annual pnl"
        #, caption =  "option prices from Delta-Neutral"

    )


```



#### Delta-Hedged PNL by Month
Let's take a closer look at the monthly PNL of the delta-hedged strategy.  Notice that significant losses occurred on August 2015, and in the first quarter of 2018.  Other than that, things look pretty good.  

This suggests an opportunity for performance enhancement via stop-loss mechanisms.


```{r, pnl_by_month, echo = FALSE, fig.width=5, fig.asp=2, fig.align="center"}

df_pnl_gathered %>%
    filter(data_date > "2014-01-01") %>% 
    filter(strategy == "delta_hedge") %>% 
    mutate(
        year = factor(lubridate::year(data_date))
        , month = lubridate::month(data_date, label = TRUE)
    ) %>% 
    group_by(year, month) %>%
    summarize(
        pnl = sum(daily_pnl)
    ) %>% 
    #ungroup() %>%
    ggplot(aes(month, pnl, fill = pnl)) +
    geom_bar(stat = "identity") +
    facet_wrap(~year, nrow = 5) +
    labs(
        title = "PNL by Month"
        , subtitle = "delta-hedged strategy"
        , x = "trade date"
        , y = "annual pnl"
        #, caption =  "option prices from Delta-Neutral"
    )
```



#### Next Steps
My initial delivery date was set for December 31, 2018.  Unfortunately, I'm simply not going to be able to hit that deadline.  As usual, all data-related projects take so much longer than I anticipate.  

I now have five years of historical data in my database, which is what allowed me to produce this report.  Additionally, I moved a lot of my backtesting code into a proper R package.  This took quite a bit of time, and learning, but was necessary to make sure that my analysis code stays organized and is scalable.  In any case, my apologies for the delay.

Here are the next steps for us to work on:

1. I am in the process of using the `rugarch` package to generate GARCH(1,1) estimates for future volatility.

1. You are working the local min/max determination.

1. Based on the two steps above, we will have an execution decision for each expiration date.  This execution decision will consist of what to sell (put, strangle, call), and what day to sell it.

1. Once these first three are complete, we can generate the PNLs for the unhedged positions (or daily delta-hedged positions).  An immediate sanity check we can do is to make sure that these daily/unhedged pnls perform better than the naive strategy outlined in this current report.  If there isn't an improvement in performance, then our "smarter decision making" isn't really helping.

1. Next, we will define the risk-management portion of the strategy.  This involves some form of delta-hedging, and a stop-loss mechanism for all the trades under extreme circumstances.

1. All of the above can be calculated in absolute dollar terms, and along the way we will be making rough estimates on capital usage.  The final stages is to figure out exactly how much capital it will take to fund this strategy, so we can know precise return numbers.


**Revised Delivery Date:** Friday, January 18th.


<!-- OLD CODE -->
```{r, calc_performance, include = FALSE}
##############
## OLD CODE ##
##############
# 
# # daily pnl for each strategy, all underlyings combined
# df_dly_pnl <-
#     df_pnl %>% 
#         group_by(frequency, data_date) %>%
#         arrange(frequency, data_date) %>% 
#         summarize(
#             naked_dly = sum(naked_dly)
#             , dh_dly = sum(dh_dly) 
#         )
# 
# 
# # total PNL for both strategies
# tot_unhedged <- df_dly_pnl$naked_dly %>% sum() %>% round(2)
# tot_hedged <- df_dly_pnl$dh_dly %>% sum() %>% round(2)
# tot_spy <- df_spy_px$pnl %>% sum(na.rm = TRUE) %>% round(2)
# 
# 
# # mean
# avg_unhedged <- df_dly_pnl$naked_dly %>% mean() %>% round(2)
# avg_hedged <- df_dly_pnl$dh_dly %>% mean() %>% round(2)
# avg_spy <- df_spy_px$pnl %>% mean(na.rm = TRUE) %>% round(2)
# 
# # standard deviation
# sd_unhedged <- df_dly_pnl$naked_dly %>% sd() %>% round(2)
# sd_hedged <- df_dly_pnl$dh_dly %>% sd() %>% round(2)
# sd_spy <- df_spy_px$pnl %>% sd(na.rm = TRUE) %>% round(2)
# 
# 
# # sharpe-ratio
# sr_unhedged <- 
#     (mean(df_dly_pnl$naked_dly) / sd(df_dly_pnl$naked_dly)) * sqrt(252) %>% round(2) # naked
# sr_hedged <- 
#     (mean(df_dly_pnl$dh_dly) / sd(df_dly_pnl$dh_dly)) * sqrt(252)  %>% round(2)# hedged
# sr_spy <- 
# (mean(df_spy_px$pnl, na.rm = TRUE) /
#         sd(df_spy_px$pnl, na.rm = TRUE)) * sqrt(252)  %>% round(2) # SPY
# 
# # returns
# ret_unhedged <- (101.27/100)^(365/581) - 1 # naked
# ret_hedged <- (103.22/100)^(365/581) - 1 # delta-hedged
# ret_spy <- (df_spy_px$adjusted[nrow(df_spy_px)]
#     /df_spy_px$adjusted[1]) ^ (365/581) - 1 # SPY 
# 
# # DVAR
# dvar_hedged <- quantile(df_dly_pnl$naked_dly, 0.05) %>% round(2)
# dvar_unhedged <- quantile(df_dly_pnl$dh_dly, 0.05) %>% round(2)
# dvar_spy <- quantile(df_spy_px$pnl, 0.05, na.rm = TRUE) %>% round(2)
# 
# df_performance <-
#     tibble(
#         strategy = c("unhedged", "hedged", "spy")
#         , total = c(tot_unhedged, tot_hedged, tot_spy)
#         , mean = c(avg_unhedged, avg_hedged, avg_spy)
#         , sd = c(sd_unhedged, sd_hedged, sd_spy)
#         , sharpe = c(sr_unhedged, sr_hedged, sr_spy)
#         , ret = c(ret_unhedged, ret_hedged, ret_spy)
#         , dvar = c(dvar_hedged, dvar_unhedged, dvar_spy)
#     )
```

```{r, testing, include = FALSE}
## exploring the largest losses
# df_pnl_gathered %>%
#     #filter(data_date > "2014-01-01") %>% 
#     filter(strategy == "delta_hedge") %>% 
#     mutate(
#         year = factor(lubridate::year(data_date))
#         , month = lubridate::month(data_date, label = TRUE)
#     ) %>% 
#     group_by(year, month) %>%
#     summarize(
#         pnl = sum(daily_pnl)
#     ) %>% 
#     filter(pnl < 0) %>% 
#     arrange(pnl) %>% 
#     ungroup() %>% 
#     top_n(-4, pnl) %>% 
#     .$pnl %>% 
#     sum()
```


<!-- #### Proposal -->
<!-- ODS would be pleased to execute a rigorous backtest analysis to inform the implementation of your trading strategy.  Of course, we will be in close contact throughout the duration of the project to ensure that you are satisfied. -->

<!-- The backtest analysis with meet the following specifications: -->

<!-- 1. Dates: December 2013 - December 2018 (5 Years)  -->

<!-- 2. Underlying: SPY (ETF) -->

<!-- 3. Approximate Trading Strategy Parameters (from your e-mail): -->
<!--     - Identify Local minima/ maxima of SPY (use 2-day RSI and VIX) -->
<!--     - Use GARCH (1,1) to determine Buy/Sell weekly 5 delta options on Monday or Tuesday prior to expiration (2 std deviation options) with bias toward selling (20,000 contracts per million) or 40% margin to equity -->
<!--     - Hedge when loss is -.5% of NAV using underlying Neutralize delta and gamma risk when portfolio hits 3% draw down -->
<!--     -  Liquidate positions when portfolio hits 10% draw down -->
<!--     - Sell puts at local bottom, sell calls at local top, sell strangles during range bound -->
<!--     - There is a distinct draw down limit per position. -->
<!--     - Hedges by selling/buying SPY futures and buying puts/calls. -->
<!-- \vspace{10pt} -->
<!-- 4. Delivery Date: 12/31/2018 -->

<!-- 5. Deliverables: html report, all R code, all raw data, all intermediate data. -->

<!-- 6. Price: \$1,500 - paid upon completion. -->




<!-- #### How to Improve This Analysis -->

<!-- 1. Include SPY in ttd_pnl graph. -->

<!-- 1. Graph PNLs by month - this is important in the alternative investment space. -->

<!-- 1. Make max width of the writing more narrow. -->

<!-- 1. Title font is too big. -->

<!-- 1. Change table column names to be more descriptive. -->

<!-- 1. Change returns to percentage. -->
