---
title: "Selling SPY Weeklys: 2014 - 2018"
subtitle: "Threshold Delta-Hedging"
author: "Pritam Dalal"
date: "February 2, 2019"
output: 
  html_document:
    theme: united
    df_print: kable
editor_options: 
  chunk_output_type: inline
---

<style>
    body .main-container {
        max-width: 700px;
    }
</style>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, load_packages, include = FALSE}
# loading packages
library(tidyverse)
library(gridExtra)
```



```{r, read_data, include=FALSE}
chr_path <- "../data_output/df_portfolio_threshold_delta_hedging_by_delta.csv"
df_portfolio <- 
    read_csv(chr_path)
```


&nbsp;

In our initial 5-year backtest of selling SPY weeklys we considered two delta-hedging strategies:

1. naked options (doing nothing)

2. rebalancing daily

In this analysis we consider an alternative approach to delta-hedging.  In particular, we set a threshold between 0% and 100%, and we rebalance the delta-hedge whenever the absolute net-delta of the position exceeds this threshold.

The the naked strategy and daily delta-hedging are two extremes of this threshold approach.  Naked option selling corresponds to a threshold of 100%; daily delta-hedging corresponds to a threshold of 0%.



```{r, calculating_metrics, include=FALSE}
df_portfolio <-
    df_portfolio %>%
        mutate(
            scaled_tot_pnl = strangle_mult * tot_pnl
        )

df_daily_pnl <-
    df_portfolio %>%
        group_by(threshold, data_date) %>%
        summarize(
            dly_tot_pnl = sum(scaled_tot_pnl)
        )


df_metrics <-
    df_daily_pnl %>%
        group_by(threshold) %>%
        summarize(
            avg_ann = (mean(dly_tot_pnl) * 252) %>% round(2)
            , std_ann = (sd(dly_tot_pnl) * sqrt(252)) %>% round(2)
            , sharpe =
                ((mean(dly_tot_pnl) / sd(dly_tot_pnl)) * sqrt(252)) %>% round(2)
            , dvar = quantile(dly_tot_pnl, 0.05)
        )

```



&nbsp;

#### Visualizing Performance as a Function of Threshold
 
In this section we calculate three key performance metrics for threshold levels ranging from 0% to 100%.  The three metrics we consider are: 

1. annualized average PNL (the numerator of the Sharpe-Ratio)

1. annualized standard deviation of PNL (the denominator of the Sharpe-Ratio)

1. Sharpe-Ratio.  

The graphs of each are given below, followed by a few words of commentary:
```{r, plot_metrics_vs_threshold, echo=FALSE}

plt_avg_ann <- 
    df_metrics %>%
        ggplot(aes(x=threshold, y=avg_ann)) +
        geom_point() +
        geom_smooth(method = 'loess', formula = 'y ~ x', se = FALSE) +
        ggtitle("Average PNL")

plt_std_ann <- 
    df_metrics %>%
        ggplot(aes(x=threshold, y=std_ann)) +
        geom_point() +
        geom_smooth(method = 'loess', formula = 'y ~ x', se = FALSE) +
        ggtitle("Standard Deviation of PNL")


plt_sharpe <-
    df_metrics %>%
        ggplot(aes(x=threshold, y=sharpe)) +
        geom_point() +
        geom_smooth(method = 'loess', formula = 'y ~ x', se = FALSE) +
        ggtitle("Sharpe-Ratio")



gridExtra::grid.arrange(plt_avg_ann, plt_std_ann, plt_sharpe, ncol=2)
```

1. The impact on standard deviation makes strong intuitive sense.  As we go from more frequent (0%) to less frequent (100%) hedging, the variability of the PNL increases.

2. The impact on average returns has intuitive appeal as well.  If we hedge too frequently (approaching 0%), we lock-in losses even when the underlying is mostly range bound.  If we hedge too *in*frequently (approaching 100%), we suffer large losses when the market moves a lot.  It's not surprising that there is a sweet spot in the middle, which in our data is a threshold of 40%-50%.

3. The Sharpe-Ratio graph has an interesting feather: there is a bifurcation between thresholds in the range of 0%-50% and those in the range of 50%-100%.  The 0%-50% group has Sharpes ranging from 1.11 to 1.38, with an average of 1.26.  The 50%-100% group has Sharpes ranging from 0.47 to 0.80, with an average of 0.62.  

```{r, scratch_work, include=FALSE}
df_metrics %>% filter(threshold > 0.5) %>% .$sharpe %>% summary()


df_metrics %>% filter(threshold %in% c(0, 0.5))
```


&nbsp;

#### Dual Scatter Plots
 
In this section we look at the two scatter plots, each with a wrinkle.  Both scatter plots graph the average annual PNL against the annual standard deviation of the pnl - one pair for corresponding to each threshold. However, instead of simply plotting a dot, we instead plot the threshold amount (left graph) and the Sharpe-Ratio (right graph).  This allows us to include three dimensions of data on a 2-D plot.

Because of how the axes of the graphs are laid out, the highest Sharpe-Ratios are in the bottom right, and the lowest Sharpe-Ratios are in the upper left.
```{r, dual_scatter_plots, echo=FALSE}
plt_thresh <-
    df_metrics %>%
        ggplot(aes(x=avg_ann, y=std_ann, label = factor(threshold))) +
        geom_text(check_overlap = TRUE, size=3) +
        ggtitle("Threshold")

plt_sharpe <-
    df_metrics %>%
        ggplot(aes(x=avg_ann, y=std_ann, label = factor(sharpe))) +
        geom_text(check_overlap = TRUE, size=3) +
        ggtitle("Sharpe-Ratio")

gridExtra::grid.arrange(plt_thresh, plt_sharpe, nrow=1)
```



Based on these two graphs, it seems that a threshold of 35%-50% has the best combination of high a Sharpe-Ratio, along with strong returns.  This mirrors the *sweet spot* we observed in the previous section.

&nbsp;

#### Conclusion & Recommendation
Based on this analysis, I would recommend a 50%-threshold delta-hedging strategy.  I suspect performance would be improved if this strategy is automated so that hedging can occur intraday, in the event of large directional moves.












