---
title: "Selling SPY Weeklys: 2014 - 2018"
subtitle: "When To Unwind Positions"
author: "Pritam Dalal"
date: "February 4, 2019"
output: 
  html_document:
    theme: united
    df_print: kable
editor_options: 
  chunk_output_type: inline
---

<style>
    body .main-container {
        max-width: 700px;
    }
</style>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, load_packages, include = FALSE}
# loading packages
library(tidyverse)
library(gridExtra)
```



```{r, read_data, include=FALSE}
chr_path <- "../data_output/df_portfolio_unwind_by_pnl_loss_portfolio.csv"
df_portfolio <- 
    read_csv(chr_path)
```


&nbsp;

In our initial 5-year backtest of selling SPY weeklys, we assumed that we would hold positions to maturity.  In this analysis, we consider adding a position unwind mechanism. 

The mechanism operates at the expiration level as follows:

1. We choose a threshold, call it $t$.  

2. For a given expiration, let $p$ be the premium sold for that expiration.  When the trade-to-date losses for that expiration exceed $t * p$, we unwind the positions.

In this analysis, we consider thresholds ranging from 10% to 1000% (10x loss).  As you might imagine, a 10x loss threshold is the same as having no unwind policy, i.e. holding all positions until maturity.

We restrict this analysis to the daily delta-hedging strategy.


```{r, calculating_metrics, include=FALSE}
# calculating the scaled, managed PNLs
df_portfolio <-
    df_portfolio %>%
    mutate(
        managed_naked = strangle_mult * dly_opt_pnl * manage_mult_naked
        , managed_dh = strangle_mult * dly_tot_pnl * manage_mult_dh
    )

# aggregating pnls by day so we can calculate Sharpe-Ratio
df_daily_pnl <-
    df_portfolio %>%
    group_by(threshold, data_date) %>%
    summarize(
        dly_managed_naked = sum(managed_naked)
        , dly_managed_dh = sum(managed_dh)
    )


# naked options
df_metrics_naked <-
    df_daily_pnl %>%
    group_by(threshold) %>%
    summarize(
        avg_ann = (mean(dly_managed_naked) * 252) %>% round(2)
        , std_ann = (sd(dly_managed_naked) * sqrt(252)) %>% round(2)
        , sharpe =
            ((mean(dly_managed_naked) / sd(dly_managed_naked)) * sqrt(252)) %>% round(2)
        
    )


# delta-hedged options
df_metrics_dh <-
    df_daily_pnl %>%
    group_by(threshold) %>%
    summarize(
        avg_ann = (mean(dly_managed_dh) * 252) %>% round(2)
        , std_ann = (sd(dly_managed_dh) * sqrt(252)) %>% round(2)
        , sharpe =
            ((mean(dly_managed_dh) / sd(dly_managed_dh)) * sqrt(252)) %>% round(2)
        
    )

```



&nbsp;

#### Visualizing Performance as a Function of Threshold

In this section we calculate three key performance metrics for loss threshold levels ranging from 10% to 1000%.  The three metrics we consider are:

1. annualized average PNL (the numerator of the Sharpe-Ratio)

1. annualized standard deviation of PNL (the denominator of the Sharpe-Ratio)

1. Sharpe-Ratio.

The graphs of each are given below, followed by a few words of commentary:
```{r, plot_metrics_vs_threshold, echo=FALSE}

plt_avg_ann <-
    df_metrics_dh %>%
        ggplot(aes(x=threshold, y=avg_ann)) +
        geom_point() +
        geom_smooth(method = 'loess', formula = 'y ~ x', se = FALSE) +
        scale_x_continuous(labels = scales::percent) +
        ggtitle("Average PNL")

plt_std_ann<-
    df_metrics_dh %>%
        ggplot(aes(x=threshold, y=std_ann)) +
        geom_point() +
        geom_smooth(method = 'loess', formula = 'y ~ x', se = FALSE) +
        scale_x_continuous(labels = scales::percent) +
        ggtitle("Standard Deviation of PNL")


plt_sharpe <-
    df_metrics_dh %>%
        ggplot(aes(x=threshold, y=sharpe)) +
        geom_point() +
        geom_smooth(method = 'loess', formula = 'y ~ x', se = FALSE) +
        scale_x_continuous(labels = scales::percent) +
        ggtitle("Sharpe-Ratio")



gridExtra::grid.arrange(plt_avg_ann, plt_std_ann, plt_sharpe, ncol=2)
```

1. Average PNL - a threshold of less than 125% seems to significantly depress returns.

2. Standard Deviation - there is a pop in variability around 125%, but this effect levels off around 250%

3. The Sharpe-Ratio ratio graph has a sweet-spot for a threshold in the range of 110%-135%.

&nbsp;

#### Dual Scatter Plots

In this section we look at the two scatter plots, each with a wrinkle.  Both scatter plots graph the average annual PNL against the annual standard deviation of the pnl - one pair for corresponding to each threshold. However, instead of simply plotting a dot, we instead plot the threshold amount (left graph) and the Sharpe-Ratio (right graph).  This allows us to include three dimensions of data on a 2-D plot.

Because of how the axes of the graphs are laid out, the highest Sharpe-Ratios are in the bottom right, and the lowest Sharpe-Ratios are in the upper left.
```{r, dual_scatter_plots, echo=FALSE}
plt_thresh <-
    df_metrics_dh %>%
    ggplot(aes(x=avg_ann, y=std_ann, label = factor(threshold))) +
    geom_text(check_overlap = TRUE, size=3) +
    ggtitle("Threshold")


# plt_thresh <-
#     df_metrics_dh %>%
#     ggplot(aes(x=avg_ann, y=std_ann, label = factor(paste0(as.character(df_metrics_dh$threshold * 100),"%")))) +
#     geom_text(check_overlap = TRUE, size=2) +
#     ggtitle("Threshold")


plt_sharpe <-
    df_metrics_dh %>%
    ggplot(aes(x=avg_ann, y=std_ann, label = factor(sharpe))) +
    geom_text(check_overlap = TRUE, size=3) +
    ggtitle("Sharpe-Ratio")

gridExtra::grid.arrange(plt_thresh, plt_sharpe, nrow=1)
```



Based on these two graphs, it seems that a threshold of 100%-130% has the highest Sharpe-Ratios.  This mirrors the *sweet spot* we observed in the previous section.

&nbsp;

#### Conclusion & Recommendation
Based on this analysis, under a daily delta-hedging regiment, I would recommend unwinding positions when losses reach around 125% of premiums sold.

It will be interesting to combine this unwind strategy with the threshold delta-hedging approach, to see if there is an optimal combination of delta-hedging threshold and position unwind threshold.

I will work on the combined analysis next.












