---
title: "Selling SPY Weeklys: 2014 - 2018"
subtitle: "Combining Risk Management Strategies"
author: "Pritam Dalal"
date: "February 4, 2019"
output: 
  html_document:
    theme: united
    df_print: kable
editor_options: 
  chunk_output_type: inline
---

<style>
    body .main-container {
        max-width: 700px;
    }
</style>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, load_packages, include = FALSE}
# loading packages
library(tidyverse)
library(gridExtra)
```



```{r, read_data, include=FALSE}
chr_path <- "../data_output/combined_risk_management_metric.csv"
df_metric <-
    read_csv(chr_path)
```


&nbsp;  

In two previous analyses, we explored two different risk management strategies to employ while selling weekly SPY options:

1. threshold delta-hedging

1. unwinding positions when a certain loss-trigger is hit

When considered in isolation, the optimal threshold for delta-hedging is in the 35%-50% range.  When daily delta-hedging, the optimal loss-trigger level for unwinding positions is around 125% of premium sold (max profits). 

In this analysis, we combine these two risk management strategies to see if there is an optimal combination of delta-hedging threshold and loss-trigger percentage.


&nbsp;

#### Sharpe Ratio Heat Map

The data is most easily visualized with a heat map of Sharpe-Ratios associated with each comination of delta-hedge threshold and loss-trigger.  We consder thresholds ranging from 0% to 100% in increments of 1%.  We consider loss triggers ranging from 10% to 300% in increments of 10%. 

The delta-hedge thresholds vary along the x-axis, increasing from left to right. The loss-trigger levels vary along the y-axis, increasing from bottom to top.

```{r, heat_map_sharpe, echo=FALSE}

df_metric %>% 
    ggplot(aes(dh_threshold, loss_trigger, label = sharpe)) +
    geom_raster(aes(fill = sharpe), interpolate = TRUE)  +
    geom_text(check_overlap = FALSE, size=2) + 
    ggtitle("Sharpe-Ratio")
```


**Observations**

1. The left-most vertical strip of the heat map represents daily delta-hedging for various loss-trigger levels.  The right-most vertical strip represents naked option selling for various loss-trigger levels.

1. The loss-trigger level has far less impact on daily delta-hedging than it does on naked option selling.  This makes sense.  Daily delta-hedging already provides a lot of risk management benefits, so overlaying an unwind strategy doesn't add that much.

1. We have enough data that I buy the overall trends, but we don't have enough data for me to believe some of the details.  For example, for infrequent delta-hedging (threshold: 50%-100%) there is a big drop off in Sharpe-Ratio from a loss trigger of 2.1 to 2.2.  The Sharpe decreases from about 1.53 to 1.08.  I suspect this is just an artifact of the market path: by increasing the trigger from 2.1 to 2.2, we just happen expose ourselves to big losses for one particular expiration.


1. The overall trend that I see - most visible in the cases of infrequent hedging - is that both low loss-triggers and high loss-triggers lead to poor Sharpe-Ratios.  This makes sense.  If the loss-trigger is too low, you prematurely unwind trades the will end up being profitable.  If the loss-trigger is too high, you expose yourself to large losses due to big moves in the underlying. 

&nbsp;

#### Conclusion

To my eye, for infrequent delta-hedging, say thresholds greater that 35%, there is a compelling case to be made for unwinding positions when losses reach about 150%-200% of max profit.  

For delta-hedging that's more frequent, it's hard to say that unwinding positions results in that much performance improvement.  However, it is still prudent to have such a policy, as it makes strong intuitive sense and it doesn't seem to *hurt* performance.  Under frequent delta-hedging, thresholds less than 35%, I would suggest a loss trigger between 100%-150% of max profit.




```{r, dual_scatter_plots, echo=FALSE}
# plt_thresh <-
#     df_metrics_dh %>%
#     ggplot(aes(x=avg_ann, y=std_ann, label = factor(threshold))) +
#     geom_text(check_overlap = TRUE, size=3) +
#     ggtitle("Threshold")


# plt_thresh <-
#     df_metrics_dh %>%
#     ggplot(aes(x=avg_ann, y=std_ann, label = factor(paste0(as.character(df_metrics_dh$threshold * 100),"%")))) +
#     geom_text(check_overlap = TRUE, size=2) +
#     ggtitle("Threshold")


# plt_sharpe <-
#     df_metrics_dh %>%
#     ggplot(aes(x=avg_ann, y=std_ann, label = factor(sharpe))) +
#     geom_text(check_overlap = TRUE, size=3) +
#     ggtitle("Sharpe-Ratio")
# 
# gridExtra::grid.arrange(plt_thresh, plt_sharpe, nrow=1)
```


&nbsp;

#### Appendix: Additional Heat Maps

Here are the heat maps for the annualized PNL (denominator of Sharpe) and the annualized standard deviation (numerator of Sharpe).

```{r, heat_map_appendix, echo=FALSE}

df_metric %>% 
    ggplot(aes(dh_threshold, loss_trigger, label = avg_ann)) +
    geom_raster(aes(fill = avg_ann), interpolate = TRUE)  +
    geom_text(check_overlap = FALSE, size=2) + 
    ggtitle("Annualized Average Daily PNL")


df_metric %>% 
    ggplot(aes(dh_threshold, loss_trigger, label = std_ann)) +
    geom_raster(aes(fill = std_ann), interpolate = TRUE)  +
    geom_text(check_overlap = FALSE, size=2) + 
    ggtitle("Annualized Standard Deviation of Daily PNL")
```









