---
title: "Selling SPY Weeklys: 2014 - 2018"
subtitle: "No-Trade Signals"
author: "Pritam Dalal"
date: "February 7, 2019"
output: 
  html_document:
    theme: united
    df_print: kable
editor_options: 
  chunk_output_type: inline
---

<style>
    body .main-container {
        max-width: 700px;
    }
</style>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, load_packages, include = FALSE}
# loading packages
library(tidyverse)
library(lubridate)
```



```{r, read_data, include=FALSE}
df_managed_pnl <- 
    read_csv("../data_output/spy_weekly_managed_pnl_10_300_5yr.csv")
df_scaling <- 
    read_csv("../data_output/spy_weekly_position_scaling_5yr.csv")
df_chain_hist <- 
    read_csv("../data_output/spy_weekly_chain_hist_garch_5yr.csv")
df_chain_desc <-
    read_csv("../data_output/spy_weekly_chain_desc_5yr.csv")


```



```{r, calc_sitout, include=FALSE}
# determining the premium by expiration
df_premium <- 
    df_scaling %>%
        filter(variation == 0.1) %>% 
        select(expiration, premium = strangle_prem_sold)


# here is the data that is used for the sit-out strategy
df_sit_out <- 
    df_managed_pnl %>% 
        filter(strategy == "strangle") %>% 
        group_by(expiration) %>% 
        summarize(
            exp_pnl = sum(scaled_managed_pnl)
            , unwind = sum(breach)
        ) %>%
        left_join(
            df_premium
            , by = "expiration"
        ) %>% 
        select(expiration, premium, exp_pnl, unwind) %>% 
        mutate(pnl_ratio = exp_pnl / premium)


# calculating sit-outs
df_sit_out$sitout <- NA
df_sit_out$sitout[1] <- FALSE
for (ix in 2:nrow(df_sit_out)){
    dbl_prev_ratio <- df_sit_out$pnl_ratio[ix - 1]
    if (dbl_prev_ratio < -2){
        df_sit_out$sitout[ix] <- TRUE
    } else {
        df_sit_out$sitout[ix] <- FALSE
    }
}



# isolating strangle pnl only
df_strangle_pnl <-
    df_managed_pnl %>% 
        filter(strategy == "strangle")

# joining on the sitout mult and calculatint the situout pnl
df_strangle_pnl <- 
    df_strangle_pnl %>% 
        left_join(
            df_sit_out %>% select(expiration, sitout)
            , by = "expiration"
        ) %>% 
        mutate(
            sitout_mult = as.integer(!sitout)
        ) %>% 
        mutate(
            sitout_pnl = scaled_managed_pnl * sitout_mult
        )


# calculating the daily pnl and the ttd pnl of the full strategy
df_sitout_daily <- 
    df_strangle_pnl %>% 
        group_by(data_date) %>% 
        summarize(
            dly_pnl = sum(scaled_managed_pnl)
            , dly_pnl_sitout = sum(sitout_pnl)
        ) %>% 
        mutate(
            ttd_pnl = cumsum(dly_pnl)
            , ttd_pnl_sitout = cumsum(dly_pnl_sitout)
        )


# gathering the columns so that we can utilize group_bys for comparison
df_sitout_comparison <- 
    df_sitout_daily %>% 
        select(data_date, daily = dly_pnl, ttd = ttd_pnl) %>% 
        mutate(signal = "base") %>% 
        bind_rows(
            df_sitout_daily %>% 
            select(data_date, daily = dly_pnl_sitout, ttd = ttd_pnl_sitout) %>% 
            mutate(signal = "sitout")
        )


# calculating performance by year
df_sitout_by_year <- 
    df_sitout_comparison %>% 
        mutate(year = year(data_date)) %>% 
        group_by(signal, year) %>% 
        summarize(
            pnl = sum(daily)
            , sharpe = (mean(daily) / sd(daily)) * sqrt(252)
        )
```



```{r, calc_forecast, include=FALSE}
# calculating the volatility premium
df_vol_premium <- 
    df_chain_desc %>% 
        select(expiration, execution) %>% 
        left_join(
            df_chain_hist
            , by = c("expiration", "execution"="trade_date")
        ) %>% 
        select(expiration, execution, bid_swap_rate, garch_forecast) %>% 
        mutate(
            vol_prem = bid_swap_rate - garch_forecast
        )


# in-sample first year mean
dbl_first_year_avg <- 
    df_vol_premium %>%  
        filter(expiration < "2015-01-01") %>% 
        .$vol_prem %>% 
        mean()

# in-sample first year standard deviation
dbl_first_year_sd <- 
    df_vol_premium %>%  
    filter(expiration < "2015-01-01") %>% 
    .$vol_prem %>% 
    sd()

# calculating z-score of volatility premium forecast
df_vol_premium$z_score <- NA_real_
df_vol_premium$garch_sell <- NA_real_
for (ix_exp in 1:nrow(df_vol_premium)){

    dt_expiration <- df_vol_premium$expiration[ix_exp]
    dbl_vol_prem <-  df_vol_premium$vol_prem[ix_exp]
        
    if (dt_expiration < "2015-01-01"){
        # for the first year, just use the in-sample number
        dbl_vol_prem_mean <- dbl_first_year_avg
        dbl_vol_prem_sd <- dbl_first_year_sd
    } else {
        # after the first year, us a trailing one year calc
        dbl_vol_prem_mean <- 
            df_vol_premium[(ix_exp - 51):ix_exp, ]$vol_prem %>% mean()
        dbl_vol_prem_sd <- 
            df_vol_premium[(ix_exp - 51):ix_exp, ]$vol_prem %>% sd()
    }
    
    # calculatin z-score of vol premium forecast
    dbl_z_score = (dbl_vol_prem - dbl_vol_prem_mean) / dbl_vol_prem_sd
    df_vol_premium$z_score[ix_exp] <- dbl_z_score
    
    
    bln_sell <- TRUE
    if (dbl_z_score < - 1.5) {
        bln_sell <- FALSE
    }
    df_vol_premium$garch_sell[ix_exp] <- bln_sell
    
}


# isolating pnl for strangle onsle onley
df_strangle_pnl <-
    df_managed_pnl %>% 
    filter(strategy == "strangle")


# joining in the volatility premium forcasting
df_strangle_pnl <- 
    df_strangle_pnl %>% 
    left_join(
        df_vol_premium %>% select(expiration, garch_sell)
        , by = "expiration"
    ) 



# calculating the pnl accoutng for the garch_based sell signals
df_strangle_pnl <- 
    df_strangle_pnl %>% 
    mutate(
        garch_pnl = scaled_managed_pnl * garch_sell
    )


# calculating the daily pnls 
df_garch_daily <- 
    df_strangle_pnl %>% 
    group_by(data_date) %>% 
    summarize(
        dly_pnl = sum(scaled_managed_pnl)
        , dly_pnl_garch = sum(garch_pnl)
    ) %>% 
    mutate(
        ttd_pnl = cumsum(dly_pnl)
        , ttd_pnl_garch = cumsum(dly_pnl_garch)
    )


# gathering a dataframe for easy comparison using group_bys
df_garch_comparison <- 
    df_garch_daily %>% 
    select(data_date, daily = dly_pnl, ttd = ttd_pnl) %>% 
    mutate(signal = "base") %>% 
    bind_rows(
        df_garch_daily %>% 
            select(data_date, daily = dly_pnl_garch, ttd = ttd_pnl_garch) %>% 
            mutate(signal = "forecast")
    )


# calculating performance by year
df_garch_by_year <- 
    df_garch_comparison %>% 
    mutate(year = year(data_date)) %>% 
    group_by(signal, year) %>% 
    summarize(
        pnl = sum(daily)
        , sharpe = (mean(daily) / sd(daily)) * sqrt(252)
    )

```




&nbsp;

The purpose of this analysis is to compare the performance impact of two different no-trade signals:



1. `simple sit-out` - sit out if there was a loss of 200% of premium sold for the previous expiration. 

1. `volatiltiy premium forecasts` - we calculate a volatility premium forecast by calculating a GARCH(1,1) forecast for realized volatility.  We then calculate the z-score of that volatility forecast.  If the z-score is less that -1.5, we don't trade.


These mechanisms are applied to the `base case` of strangle selling with 10% threshold delta-hedging, and 300% loss-trigger position unwinds.

&nbsp;



#### Overall Peformance
The `forecast` mechanism hurts performance slightly, while the `sit-out` mechanism significantly enhances performance.

```{r, overall_performance, echo=FALSE}
df_signal_comparison <- 
    df_sitout_comparison %>% 
        bind_rows(
            df_garch_comparison %>% filter(signal == "forecast")
        )

df_signal_comparison %>%
    group_by(signal) %>%
    summarize(
        total_pnl = sum(daily) %>% round(2)
        , stdev_pnl = (sd(daily) * sqrt(252)) %>% round(2)
        , sharpe = ((mean(daily) / sd(daily)) * sqrt(252)) %>% round(2)
    )

```



&nbsp;

#### Visualizing Cumulative Performance

By observing the cumulative performance graph it's easy to see that most of the out-performance of the `sit-out` signal comes from avoding an extremely costly expiration in early 2018.  






```{r, ttd_pnl_graph, echo=FALSE}

df_signal_comparison %>%
    ggplot(aes(x = data_date, y = ttd, color = signal)) +
    geom_line() +
    labs(
        title = "Cumulative Performance"
        , subtitle = "no-trade signals"
        , x = "trade date"
        , y = "ttd pnl"
    )
```


Both  signals are a slight drag on performance in the good times, which should be expected.  By employing a no-trade signals, you will no doubt lose out on some false-positive winners.  But the question is whether that inevitable cost will come with the benefit of avoiding some painful trades.  The answer is yes for `sit-out` but not for `forecast`.


&nbsp;

#### Performance by Year

It is instructive to analyze performance of each strategy by year.  The `sit-out` has a negative performance impact in all years except for 2018, but the savings in 2018 are substantial.


<!-- &nbsp; -->
```{r, performanc_by_year, echo=FALSE}
df_performance_by_year <-
    df_signal_comparison %>%
        mutate(year = year(data_date)) %>%
        group_by(signal, year) %>%
        summarize(
            pnl = sum(daily)
            , sharpe = (mean(daily) / sd(daily)) * sqrt(252)
        )

# pnl by year
df_performance_by_year %>%
    filter(year > 2013) %>%
    ggplot(aes(factor(year), pnl, fill = signal)) +
        geom_bar(stat = "identity", position = "dodge") +
        labs(
            title = "PNL by Year"
            , subtitle = "no-trade signals"
            , x = "trade date"
            , y = "annual pnl"
        )

# pnl by year
df_performance_by_year %>%
    filter(year > 2013) %>%
    ggplot(aes(factor(year), sharpe, fill = signal)) +
        geom_bar(stat = "identity", position = "dodge") +
        labs(
            title = "Sharpe by Year"
            , subtitle = "strangles vs puts"
            , x = "trade date"
            , y = "annual sharpe"
        )
```



&nbsp;

#### Conclusion & Recommendation

Selling options is good until it isn't.  The key to a successful option selling is proper risk management.  The simple strategy of sitting out a week after a significant loss makes strong intuitive sense, and its success comes as no suprise given the well known phenomenon of realized volatility clustering.

Now it's true that the success of the `sit-out` signal is largely based on avoiding certain trades in 2018, but it really doesn't hurt much in other years (the good times).

When the markets are experience significant turmoil, it's not a bad idea to sit out for a bit.  There will be plenty of time to make money again when turbulence subsides.

My next analysis will assume the employment of `sit-out` no-trade signal.







